How to compute the fusion product of MV cycles in type A

### 2021-04-01 draft.tex
added more detail to ES fact

\begin{itemize}
    \item For $ s \in \CC$ 
    % and $\lambda\in P_{++}$
    we have a Jordan form matrix $ J_{s,\lambda}$ with eigenvalue $ s$ and Jordan blocks of sizes $ \lambda_1, \dots, \lambda_m$. 
    % \acom{$s\in\AA$?}
    % 
    \item The adjoint orbit $ \OO^\lambda \subset M_N(\CC)$ of matrices conjugate to $ J_{0,\lambda}$.  Its closure is $ \overline{\OO}^\lambda = \bigcup_{\lambda' \le \lambda} \OO^{\lambda'}$.
    % 
    \item For $ s \in \CC, s \ne 0$, the adjoint orbit $ \OO^{\lambda_1, \lambda_2}_{0,s}$ of matrices conjugate to $ J_{0,\lambda_1} \oplus J_{s,\lambda_2}$.  Its closure is $ \overline{\OO}^{\lambda_1, \lambda_2}_{0,s} = \bigcup_{\lambda_1' \le \lambda_1, \lambda'_2 \le \lambda_2} \OO^{\lambda_1', \lambda'_2}_{0,s}$. 
    % \acom{$s\in\AA$?}
    % 
    \item A linear operator $ T : V \rightarrow V $ on an $N$-dimensional vector space $V$ is said to have Jordan type $((0,\lambda_1), (s,\lambda_2))$ if its matrix representatives lie in $ \OO^{\lambda_1, \lambda_2}_{0,s}$.
    % 
    \item The family of adjoint orbits $ \overline{\OO}^{\lambda_1, \lambda_2}_{0,\AA} \rightarrow \AA$ whose fibre over $ s \ne 0 $ is $ \overline{\OO}^{\lambda_1, \lambda_2}_{0,s}$ and whose fibre over 0 is $\overline{\OO}^\lambda$.

@@

\begin{proposition}
    The family $\OO^{\lambda_1,\lambda_2}_{0,\AA}$ defined by \Cref{above} has zero fibre $\OO^\lambda$. 
\end{proposition}

We will show that this is a corollary of the following result on ``rank varieties''.

\begin{theorem}(\cite{eisenbud1989rank})
Let $r$ be a decreasing concave non-negative integral function with $r(0) = N$ (the dimension of the ambient vector space). Such a function is called a rank function. Then, in the space of $N\times N$ matrices, 
% The relevant result from Eisenbud--Saltman seems to be that 
$$X_r = X_{r,0} = \{A:\rk A^i \le r(i) \}$$ 
is the flat limit of 
% $$X_{r,\lambda} = \{A : \rk (A-\lambda_i)^j\le r(i,j)\}$$
$$X_{r,z} = \{A : \rk (A-z_i)^j\le r(i,j)\}$$
with $r(i,j)$ determined by $r(i)$.  
\end{theorem}

The case that we are interested in is the case of two eigenvalues: $(z_1,z_2) = (0,s)$. \acom{Does order matter?} 
% Let me change notation so that $\lambda_i$ are shapes and 
We begin by considering the open loci where the rank conditions are strict.

@@

Next, take $p$ to be a permutation of the (lengths of) columns $a_j$ of $\lambda$ such that $a_{p(1)},\dots,a_{p(m_1)}$ are the (lengths of) columns of $\lambda_1$ and $a_{p(m_1 + 1)},\dots, a_{p(m_1 + m_2)}$ are the (lengths of) columns of $\lambda_2$. 
Then the functions 
$$
\begin{aligned}
r(1,j) &= N - a_{p(1)} - \cdots - a_{p(j)} \\
% \text{ and } 
r(2,j) &= N - a_{p(m_1 + 1)} - \cdots - a_{p(m_1 + j)}
\end{aligned}
$$
are such that 
\begin{align*}
    \OO^{\lambda_1,\lambda_2}_{0,s} = X_{r,(0,s)} = \{A:\rk A^j &= r(1,j) \text{ for } i = 1,\dots, m_1 \\ \text{ and } \rk (A-s)^j &= r(2,j) \text{ for } i = m_1 + 1 ,\dots, m \}
\end{align*}
and according to \cite[Theorem 2.1(iii)]{eisenbud1989rank} the family $X_{r,\AA^m}$ is flat over $\AA^m$ with zero fibre $X_{r,0} = \OO^\lambda$ as desired. I guess we need to carefully choose the $W \subset \AA^m$ defining our family, which is smaller than the family $X_{r,\AA^m}$. Well, $W$ is just cut out by $0 = x_1 = x_2 = \cdots = x_{m_1}$ and $ x_{m_1 + 1} = \cdots = x_m$. Or some permutation thereof to match the permutation $p$ above.

@@

    Fix $i \in \{1,\dots,m\}$ and let $p = \mu^1 + \cdots + \mu^i$. 
    % Define $\lambda^{(i)}(\tau_1 \circ \tau_2)$ so that the 
    Jordan type of $t\big|_{\CC[t]^p/L(0)\cap\CC[t]^p}$ is $\lambda_1^{(i)}$ and Jordan type of $t\big|_{\CC[t]^p/L(s)\cap \CC[t]^p}$ is $\lambda_2^{(i)}$.

>>>>

\begin{itemize}
    \item For $ s \in \CC$ 
    % and $\lambda\in P_{++}$
    we have a Jordan form matrix $ J_{s,\lambda}$ with eigenvalue $ s$ and Jordan blocks of sizes $ \lambda_1, \dots, \lambda_m$. 
    % \acom{$s\in\AA$?}
    % 
    \item The adjoint orbit $ \OO^\lambda \subset M_N(\CC)$ of matrices conjugate to $ J_{0,\lambda}$.  Its closure is $ \overline{\OO}^\lambda = \bigcup_{\lambda' \le \lambda} \OO^{\lambda'}$.
    % 
    \item For $ s \in \CC, s \ne 0$, the adjoint orbit $ \OO^{\lambda_1, \lambda_2}_{0,s}$ of matrices conjugate to $ J_{0,\lambda_1} \oplus J_{s,\lambda_2}$.  Its closure is $ \overline{\OO}^{\lambda_1, \lambda_2}_{0,s} = \bigcup_{\lambda_1' \le \lambda_1, \lambda'_2 \le \lambda_2} \OO^{\lambda_1', \lambda'_2}_{0,s}$. 
    % \acom{$s\in\AA$?}
    % 
    \item A linear operator $ T : V \rightarrow V $ on an $N$-dimensional vector space $V$ is said to have Jordan type $((0,\lambda_1), (s,\lambda_2))$ if its matrix representatives lie in $ \OO^{\lambda_1, \lambda_2}_{0,s}$.
    % 
    \item The family of adjoint orbits $ \overline{\OO}^{\lambda_1, \lambda_2}_{0,\AA} \rightarrow \AA$ whose fibre over $ s \ne 0 $ is $ \overline{\OO}^{\lambda_1, \lambda_2}_{0,s}$ and whose fibre over 0 is $\overline{\OO}^\lambda$ a consequence of Eisenbud and Salten's description/characterization(?) of ``rank varieties'' as proved in \Cref{cor:es}.

@@

\begin{theorem}(\cite{eisenbud1989rank})
    \label{thm:es}
Let $r$ be a decreasing concave non-negative integral function with $r(0) = N$ (the dimension of the ambient vector space). Such a function is called a rank function. Then, in the space of $N\times N$ matrices, 
% The relevant result from Eisenbud--Saltman seems to be that 
$$X_r = X_{r,0} = \{A:\rk A^i \le r(i) \}$$ 
is the flat limit of 
% $$X_{r,\lambda} = \{A : \rk (A-\lambda_i)^j\le r(i,j)\}$$
$$X_{r,z} = \{A : \rk (A-z_i)^j\le r(i,j)\}$$
with $r(i,j)$ determined by $r(i)$.  
\end{theorem}

\acom{Don't know how much of this setup to recall but here are the parts that will appear in the quoted theorem.}
% 
ES show that 
\begin{enumerate}
    \item $X_r$ is a normal variety
    \item $X_r$ is Gorenstein with rational singularities
    \item $X_r$ fits into a flat family over $\AA^m$ of normal varieties whose fibre over a point $(z_1,\dots,z_m)$ such that the $z_i$ are all distinct is 
\end{enumerate}
\[
    X_{r;z_1,\dots,z_m} = \left\{
        A \in \End(V) : \cork(A - z_i) \ge r(i-1) - r(i) , i = 1,\dots,m 
        \right\}
\]
They construct a simultaneous resolution of singularities and use the resolution of $X_r$ to describe the tangent space to $X_r$ at a point corresponding to an endomorphism $A$ such that $\rk A^i = r_i$ for all $i$. As an application of the tangent space computation they show that the general fiber of this family can be written as the scheme-theoretic intersection of varieties like $X_r$.

To construct the resolution, they take $m$ to be the largest number such that $r(m) \ne r(m+1)$.
% Fact: rank functions stabilize
Let $(x_1,\dots, x_m)$ be coordinates on $\AA^m$ and let $W\subset \AA^m$ be a linear subvariety (translate of subvectorspace). 
They introduce the flag variety $F = \{(V = V_0 \supset V_1 \supset \cdots \supset V_m \supset 0 ) : \dim V_i = r(i)\}$ and define $\mathcal X_{r,W}\subset\End(V) \times W \times F$ to be the subvariety of triples $(A,\vec z,\{V_i\})$ such that $(A-z_i) V_{i-1} \subset V_i$ for $z_i = x_i(\vec z)$ and $i = 1,\dots,m$. The (reduced image of) projection of $\mathcal X_{r,W}$ down to $\End(V) \times W$ is denoted $X_{r,W}$ a closed affine subvariety.

\begin{theorem}
    (\cite[Theorem~2.1 iii)]{eisenbud1989rank})
    $X_{r,W}$ is the restriction of $X_{r,\AA^m}$ to $W$ and it is flat over $W$. 
\end{theorem}

They apply this result to study the fibers of the family $X_{r,\AA^m}$ over points $\vec z \in \AA^m$ for $z_i$ not necessarily distinct. They partition the coordinates $z_i$ of $\vec z$ by equality with the help of a permutation $p$ of $\{1,\dots,m\}$ such that 
\begin{align*}
    z_{p(1)} &= \cdots = z_{p(m_1)} , \\
    z_{p(m_1 + 1)} &= \cdots = z_{p(m_2)} , \\ 
    \dots \\
    z_{p(m_s + 1)} &= \cdots = z_{p(m_s + 1)}
\end{align*}
and $p$ preserves the order within each interval $m_i + 1,\dots,m_{i+1}$. \acom{Not sure what this means. Order of coordinates?} With this notation in place they define 
\begin{gather*}
    r(i,j):= \dim V - a_{p(m_i + 1)} - \cdots - a_{p(m_i + j )} \\
    \qquad \qquad \qquad \qquad \qquad \text{ for } i = 1,\dots,s \text{ and } j = 1,\dots , m_{i + 1 } - m_i 
\end{gather*}
where $a_i = r(i-1) - r(i)$ which characterizes the Jordan types of the generalized eigenspaces associated to the eigenvalues $z_i$. Finally they state the following corollary. 
\begin{corollary}
    (\cite[Corollary~2.2]{eisenbud1989rank})
    The scheme-theoretic fibre of $X_{r,W}$ over $\vec z\in W$ is 
    \[
        X_{r,\vec z} = \{A \in \End(V) : \rk (A-z_i)^j\le r(i,j) \text{ for all } i , j \}
    \]
\end{corollary}

\begin{corollary}
    \label{cor:es}
    The family $\OO^{\lambda',\lambda''}_{0,\AA}$ defined by \Cref{above} has zero fibre $\OO^\lambda$. 
\end{corollary}

\begin{lemma}
    \label{lem:es}
    Let $\lambda = \lambda' + \lambda''$ be permutations. There exists a permutation of the columns of $\lambda$ such that the first $\lambda'_1$ columns are precisely the columns of $\lambda'$ and the next $\lambda''_1$ columns are precisely the columns of $\lambda''$. 
\end{lemma}

\begin{lemma}
   $r(j) = N - \#\text{~boxes in first $j$ columns of }\lambda$ is a rank function. 
\end{lemma}

The case that we are interested in is the case of two generalized/repeated eigenvalues, i.e.\ $\vec z = (z_1,z_2) = (0,s)$. \acom{Does order matter?} 
% Let me change notation so that $\lambda_i$ are shapes and 
We begin by considering the open loci where the rank conditions are strict.

@@

Next, take $p$ to be a permutation of the (lengths of) columns $a_j$ of $\lambda$ such that $a_{p(1)},\dots,a_{p(m_1)}$ are the (lengths of) columns of $\lambda_1$ and $a_{p(m_1 + 1)},\dots, a_{p(m_1 + m_2)}$ are the (lengths of) columns of $\lambda_2$. 
Then the functions 
$$
\begin{aligned}
r(1,j) &= N - a_{p(1)} - \cdots - a_{p(j)} \\
% \text{ and } 
r(2,j) &= N - a_{p(m_1 + 1)} - \cdots - a_{p(m_1 + j)}
\end{aligned}
$$
are such that 
\begin{align*}
    \OO^{\lambda_1,\lambda_2}_{0,s} = X_{r,(0,s)} = \{A:\rk A^j &= r(1,j) \text{ for } i = 1,\dots, m_1 \\ \text{ and } \rk (A-s)^j &= r(2,j) \text{ for } i = m_1 + 1 ,\dots, m \}
\end{align*}
and according to \cite[Theorem 2.1(iii)]{eisenbud1989rank} the family $X_{r,\AA^m}$ is flat over $\AA^m$ with zero fibre $X_{r,0} = \OO^\lambda$ as desired. I guess we need to carefully choose the $W \subset \AA^m$ defining our family, which is smaller than the family $X_{r,\AA^m}$. Well, $W$ is just cut out by $0 = x_1 = x_2 = \cdots = x_{m_1}$ and $ x_{m_1 + 1} = \cdots = x_m$, or some permutation thereof to match the permutation $p$ above.

@@

    Fix $i \in \{1,\dots,m\}$ and let $p = \mu^1 + \cdots + \mu^i$. 
    % Define $\lambda^{(i)}(\tau_1 \circ \tau_2)$ so that the 
    Jordan type of $t\big|_{\CC[t]^p/L(0)\cap\CC[t]^p}$ is $\lambda_1^{(i)}$ and Jordan type of $t-s\big|_{\CC[t]^p/L(s)\cap \CC[t]^p}$ is $\lambda_2^{(i)}$.
